---
title: "Advertising Sales Prediction"
author: "Trevor Okinda"
date: "2024"
output:
  github_document: 
    toc: yes
    toc_depth: 4
    fig_width: 6
    fig_height: 4
    df_print: default
editor_options:
  chunk_output_type: console
---

# Student Details

|                                              |     |
|----------------------------------------------|-----|
| **Student ID Number**                        | 134780 |
| **Student Name**                             | Trevor Okinda |
| **BBIT 4.2 Group**                           | C |
| **Project Name**                             | Advertising Sales Prediction |

# Setup Chunk

**Note:** the following KnitR options have been set as the global defaults: <BR> `knitr::opts_chunk$set(echo = TRUE, warning = FALSE, eval = TRUE, collapse = FALSE, tidy = TRUE)`.

More KnitR options are documented here <https://bookdown.org/yihui/rmarkdown-cookbook/chunk-options.html> and here <https://yihui.org/knitr/options/>.

```{r setup, include=FALSE}
library(formatR)
knitr::opts_chunk$set(
  warning = FALSE,
  collapse = FALSE
)
```

### Source: 

The dataset that was used can be downloaded here: *\<https://www.kaggle.com/datasets/yasserh/advertising-sales-dataset\>*

### Reference:

*\<M Yasser H. (n.d.). Advertising Sales Dataset [Data set]. Kaggle. https://www.kaggle.com/datasets/yasserh/advertising-sales-dataset\>\
Refer to the APA 7th edition manual for rules on how to cite datasets: <https://apastyle.apa.org/style-grammar-guidelines/references/examples/data-set-references>*


# Understanding the Dataset (Exploratory Data Analysis (EDA))

## Loading the Dataset
```{r Loading Dataset}
# Load dataset
AdData <- read.csv("Advertising_Budget_and_Sales.csv", colClasses = c(
  TV_Ad_Budget = "numeric",
  Radio_Ad_Budget = "numeric",
  Newspaper_Ad_Budget = "numeric",
  Sales = "numeric"
))

# Display structure to verify data types
str(AdData)

# Display first few rows to ensure data is loaded correctly
head(AdData)

View(AdData)
```

## Measures of Frequency
```{r MOF}
# Frequency for each budget type and sales (bins for better insight)
hist(AdData$TV_Ad_Budget, main = "Frequency of TV Ad Budget", xlab = "TV Ad Budget", col = "lightblue")
hist(AdData$Radio_Ad_Budget, main = "Frequency of Radio Ad Budget", xlab = "Radio Ad Budget", col = "lightgreen")
hist(AdData$Newspaper_Ad_Budget, main = "Frequency of Newspaper Ad Budget", xlab = "Newspaper Ad Budget", col = "lightcoral")
hist(AdData$Sales, main = "Frequency of Sales", xlab = "Sales", col = "lightyellow")
```

## Measures of Central Tendency
```{r MOCT}
# Mean
mean_tv <- mean(AdData$TV_Ad_Budget)
mean_radio <- mean(AdData$Radio_Ad_Budget)
mean_newspaper <- mean(AdData$Newspaper_Ad_Budget)
mean_sales <- mean(AdData$Sales)

# Median
median_tv <- median(AdData$TV_Ad_Budget)
median_radio <- median(AdData$Radio_Ad_Budget)
median_newspaper <- median(AdData$Newspaper_Ad_Budget)
median_sales <- median(AdData$Sales)

# Mode function (R does not have a built-in mode function)
get_mode <- function(x) {
  uniq_vals <- unique(x)
  uniq_vals[which.max(tabulate(match(x, uniq_vals)))]
}

mode_tv <- get_mode(AdData$TV_Ad_Budget)
mode_radio <- get_mode(AdData$Radio_Ad_Budget)
mode_newspaper <- get_mode(AdData$Newspaper_Ad_Budget)
mode_sales <- get_mode(AdData$Sales)

# Display central tendency measures
list(
  Mean = c(TV = mean_tv, Radio = mean_radio, Newspaper = mean_newspaper, Sales = mean_sales),
  Median = c(TV = median_tv, Radio = median_radio, Newspaper = median_newspaper, Sales = median_sales),
  Mode = c(TV = mode_tv, Radio = mode_radio, Newspaper = mode_newspaper, Sales = mode_sales)
)
```

## Measures of Distribution
```{r Measures of Distribution}
# Standard deviation and variance
sd_tv <- sd(AdData$TV_Ad_Budget)
var_tv <- var(AdData$TV_Ad_Budget)

sd_radio <- sd(AdData$Radio_Ad_Budget)
var_radio <- var(AdData$Radio_Ad_Budget)

sd_newspaper <- sd(AdData$Newspaper_Ad_Budget)
var_newspaper <- var(AdData$Newspaper_Ad_Budget)

sd_sales <- sd(AdData$Sales)
var_sales <- var(AdData$Sales)

# Range and Interquartile Range (IQR)
range_tv <- range(AdData$TV_Ad_Budget)
iqr_tv <- IQR(AdData$TV_Ad_Budget)

range_radio <- range(AdData$Radio_Ad_Budget)
iqr_radio <- IQR(AdData$Radio_Ad_Budget)

range_newspaper <- range(AdData$Newspaper_Ad_Budget)
iqr_newspaper <- IQR(AdData$Newspaper_Ad_Budget)

range_sales <- range(AdData$Sales)
iqr_sales <- IQR(AdData$Sales)

# Display distribution measures
list(
  Standard_Deviation = c(TV = sd_tv, Radio = sd_radio, Newspaper = sd_newspaper, Sales = sd_sales),
  Variance = c(TV = var_tv, Radio = var_radio, Newspaper = var_newspaper, Sales = var_sales),
  Range = list(TV = range_tv, Radio = range_radio, Newspaper = range_newspaper, Sales = range_sales),
  IQR = c(TV = iqr_tv, Radio = iqr_radio, Newspaper = iqr_newspaper, Sales = iqr_sales)
)
```

## Measures of Relationship
```{r MOR}
# Correlation matrix
cor_matrix <- cor(AdData[, c("TV_Ad_Budget", "Radio_Ad_Budget", "Newspaper_Ad_Budget", "Sales")])

# Scatter plot matrix
pairs(AdData[, c("TV_Ad_Budget", "Radio_Ad_Budget", "Newspaper_Ad_Budget", "Sales")], main = "Scatterplot Matrix")

# Display correlation matrix
cor_matrix
```

## Plots
```{r Plots}
# Load necessary library
library(ggplot2)

# Histogram for each variable
ggplot(AdData, aes(x = TV_Ad_Budget)) +
  geom_histogram(binwidth = 20, fill = "lightblue", color = "black") +
  labs(title = "Distribution of TV Ad Budget", x = "TV Ad Budget", y = "Frequency")

ggplot(AdData, aes(x = Radio_Ad_Budget)) +
  geom_histogram(binwidth = 5, fill = "lightgreen", color = "black") +
  labs(title = "Distribution of Radio Ad Budget", x = "Radio Ad Budget", y = "Frequency")

ggplot(AdData, aes(x = Newspaper_Ad_Budget)) +
  geom_histogram(binwidth = 10, fill = "lightcoral", color = "black") +
  labs(title = "Distribution of Newspaper Ad Budget", x = "Newspaper Ad Budget", y = "Frequency")

ggplot(AdData, aes(x = Sales)) +
  geom_histogram(binwidth = 5, fill = "lightyellow", color = "black") +
  labs(title = "Distribution of Sales", x = "Sales", y = "Frequency")

# Boxplot for each variable
ggplot(AdData, aes(y = TV_Ad_Budget)) + 
  geom_boxplot(fill = "lightblue") + 
  labs(title = "Boxplot of TV Ad Budget", y = "TV Ad Budget")

ggplot(AdData, aes(y = Radio_Ad_Budget)) + 
  geom_boxplot(fill = "lightgreen") + 
  labs(title = "Boxplot of Radio Ad Budget", y = "Radio Ad Budget")

ggplot(AdData, aes(y = Newspaper_Ad_Budget)) + 
  geom_boxplot(fill = "lightcoral") + 
  labs(title = "Boxplot of Newspaper Ad Budget", y = "Newspaper Ad Budget")

ggplot(AdData, aes(y = Sales)) + 
  geom_boxplot(fill = "lightyellow") + 
  labs(title = "Boxplot of Sales", y = "Sales")

# Density plot for each variable
ggplot(AdData, aes(x = TV_Ad_Budget)) + 
  geom_density(fill = "lightblue") + 
  labs(title = "Density Plot of TV Ad Budget", x = "TV Ad Budget")

ggplot(AdData, aes(x = Radio_Ad_Budget)) + 
  geom_density(fill = "lightgreen") + 
  labs(title = "Density Plot of Radio Ad Budget", x = "Radio Ad Budget")

ggplot(AdData, aes(x = Newspaper_Ad_Budget)) + 
  geom_density(fill = "lightcoral") + 
  labs(title = "Density Plot of Newspaper Ad Budget", x = "Newspaper Ad Budget")

ggplot(AdData, aes(x = Sales)) + 
  geom_density(fill = "lightyellow") + 
  labs(title = "Density Plot of Sales", x = "Sales")

# Scatter plots with regression line
ggplot(AdData, aes(x = TV_Ad_Budget, y = Sales)) + 
  geom_point(color = "blue") + 
  geom_smooth(method = "lm", color = "red") + 
  labs(title = "TV Ad Budget vs Sales", x = "TV Ad Budget", y = "Sales")

ggplot(AdData, aes(x = Radio_Ad_Budget, y = Sales)) + 
  geom_point(color = "green") + 
  geom_smooth(method = "lm", color = "red") + 
  labs(title = "Radio Ad Budget vs Sales", x = "Radio Ad Budget", y = "Sales")

ggplot(AdData, aes(x = Newspaper_Ad_Budget, y = Sales)) + 
  geom_point(color = "coral") + 
  geom_smooth(method = "lm", color = "red") + 
  labs(title = "Newspaper Ad Budget vs Sales", x = "Newspaper Ad Budget", y = "Sales")

# Pairwise scatter plot matrix
pairs(AdData[, c("TV_Ad_Budget", "Radio_Ad_Budget", "Newspaper_Ad_Budget", "Sales")], main = "Pairwise Scatter Plot Matrix")

# Correlation heatmap
library(reshape2)
cor_data <- cor(AdData[, c("TV_Ad_Budget", "Radio_Ad_Budget", "Newspaper_Ad_Budget", "Sales")])
melted_cor_data <- melt(cor_data)

ggplot(data = melted_cor_data, aes(x = Var1, y = Var2, fill = value)) +
  geom_tile() +
  scale_fill_gradient2(low = "red", high = "blue", mid = "white", midpoint = 0, limit = c(-1,1)) +
  theme_minimal() +
  labs(title = "Correlation Heatmap", x = "", y = "") +
  geom_text(aes(label = round(value, 2)), color = "black", size = 4)


```

# Preprocessing and Data Transformation
## Missing Values
```{r Missing Values}
# Check for missing values in each column
missing_values_summary <- sapply(AdData, function(x) sum(is.na(x)))
missing_values_summary

# Total number of missing values in the dataset
total_missing_values <- sum(is.na(AdData))
total_missing_values

# Load necessary library for visualization
library(VIM)

# Visualize missing values
aggr_plot <- aggr(AdData, col = c('navyblue', 'orange'), numbers = TRUE, sortVars = TRUE, 
                  labels = names(AdData), cex.axis = 0.7, gap = 3, ylab = c("Missing data", "Pattern"))


```

# Training Model
## Data Splitting
```{r Data Splitting}
library(caret)
# Set seed for reproducibility
set.seed(123)

# Define the proportion of data to use for training
train_size <- 0.7  # 70% for training, 30% for testing

# Create a random sample of row indices for the training set
train_indices <- sample(1:nrow(AdData), size = train_size * nrow(AdData))

# Split the data into training and testing sets
train_data <- AdData[train_indices, ]
test_data <- AdData[-train_indices, ]

# Check the dimensions of the training and testing sets
dim(train_data)
dim(test_data)
```

## Bootstrapping
```{r Bootstrapping}
library(boot)
# Set seed for reproducibility
set.seed(123)

# Number of bootstrap samples
n_bootstrap <- 1000

# Initialize a vector to store bootstrap sample means
bootstrap_means <- numeric(n_bootstrap)

# Perform bootstrapping
for (i in 1:n_bootstrap) {
  # Sample with replacement
  sample_data <- AdData[sample(1:nrow(AdData), replace = TRUE), ]
  
  # Calculate the mean of Sales for the bootstrap sample
  bootstrap_means[i] <- mean(sample_data$Sales)
}

# Display the results
mean(bootstrap_means)  # Average of the bootstrap sample means
sd(bootstrap_means)    # Standard deviation of the bootstrap sample means

# Plot the bootstrap distribution
hist(bootstrap_means, main = "Bootstrap Distribution of Sales Mean",
     xlab = "Mean Sales", col = "lightblue", border = "black")
```

## Cross-validation
```{r Cross-validation}
# Load necessary libraries
library(caret)

# Set seed for reproducibility
set.seed(123)

# Define the number of folds for cross-validation
k <- 5

# Create the training control
train_control <- trainControl(method = "cv", number = k)

# Fit a linear model (you can change this to any model of your choice)
model <- train(Sales ~ TV_Ad_Budget + Radio_Ad_Budget + Newspaper_Ad_Budget, 
               data = AdData, 
               method = "lm", 
               trControl = train_control)

# Display the results
print(model)
```

## Training Different Models
```{r Training Different Models}
# Load necessary libraries
library(caret)
library(glmnet)      # For Lasso Regression
library(randomForest) # For Random Forest Regression

# Set seed for reproducibility
set.seed(123)

# Create the training control
train_control <- trainControl(method = "cv", number = 5)  # 5-fold cross-validation

# 1. Linear Regression
linear_model <- train(Sales ~ TV_Ad_Budget + Radio_Ad_Budget + Newspaper_Ad_Budget, 
                      data = AdData, 
                      method = "lm", 
                      trControl = train_control)


# 2. Random Forest Regression
rf_model <- train(Sales ~ TV_Ad_Budget + Radio_Ad_Budget + Newspaper_Ad_Budget, 
                  data = AdData, 
                  method = "rf", 
                  trControl = train_control)


# 3. Lasso Regression
lasso_model <- train(Sales ~ TV_Ad_Budget + Radio_Ad_Budget + Newspaper_Ad_Budget,
                     data = AdData,
                     method = "glmnet",
                     trControl = train_control,
                     tuneGrid = expand.grid(alpha = 1, lambda = seq(0, 0.1, by = 0.01)))

# Display results for all models
print(linear_model)
print(lasso_model)
print(rf_model)
```

## Performance Comparison
```{r Performance Comparison}
# Combine the results from the models into a resamples object
model_list <- resamples(list(Linear = linear_model, 
                             Lasso = lasso_model, 
                             RandomForest = rf_model))

# Summary of the resamples
summary(model_list)

# Visualization of the performance comparison
bwplot(model_list)  # Boxplot to compare RMSE and other metrics

```

## Saving Model
```{r Saving Model}
# Load the saved model
loaded_rf_model <- readRDS("./models/saved_rf_model.rds")

# Create new data for predictions
new_data <- data.frame(
  TV_Ad_Budget = 100,  # Example value for TV Ad Budget
  Radio_Ad_Budget = 20,  # Example value for Radio Ad Budget
  Newspaper_Ad_Budget = 30  # Example value for Newspaper Ad Budget
)

# Use the loaded model to make predictions
predictions_loaded_model <- predict(loaded_rf_model, newdata = new_data)

# Print predictions
print(predictions_loaded_model)

```

